% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Today's dominant hardware description languages (HDLs), namely Verilog and VHDL, rely on limited register-transfer-level (RTL) constructs. These constructs tightly couple design functionality with timing requirements and target-device constraints. As hardware designs and device architectures become increasingly more complex, these dominant HDLs yield verbose and unportable code.
To raise the level of abstraction, several high-level synthesis (HLS) tools were introduced, usually based on software languages such as C. Unfortunately, designing hardware with sequential software language semantics comes at a price; the designer loses the ability to control hardware construction and data scheduling, which is crucial in many design use-cases. 

In this paper we further extend DFiant, a Scala-based HDL that uses the dataflow model to decouple functionality from implementation constraints.
DFiant's frontend enables functional bit-accurate hardware description, while maintaining a complete timing-agnostic and device-agnostic code. DFiant bridges the gap between software programming and hardware construction, driving an intuitive functional object oriented code into a high-performance hardware implementation.

For a proof of concept, we implemented a compiler frontend for DFiant, which transforms DFiant code into a dataflow graph, and a preliminary auto-pipelining backend, which maps the graph into synthesizable Verilog code. We further implemented two test cases in DFiant: an Advanced Encryption Standard cipher block and an IEEE-754 floating point multiplier. We compared both test cases against modern design flows. Our results demonstrate that DFiant can greatly simplify hardware designs yet still maintain competitive performance.
\end{abstract}

%We defined a new language from the ground up, borrowing concepts from hardware, dataflow and software languages. The result is a strongly-typed, purely synthesizable extendable language frontend, fit for both general-purpose and high level hardware description.
